\documentclass{article}

\input{preamble-articles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% We define a command \@ifnextcharamong accepting an arbitrary number of
%%%% arguments. The first is what it should do if a match is found, the second
%%%% contains what it should do when no match is found; all the other arguments
%%%% are the things it tries to find as the next character.
%%%%
%%%% For example \@ifnextcharamong{#1}{#2}{*}{\bgroup} expands #1 if the next
%%%% character is a * or a \bgroup and it expands #2 otherwise.

\makeatletter
\newcommand{\@ifnextcharamong}[2]
  {\@ifnextchar\bgroup{\@@ifnextchar{#1}{\@@ifnextcharamong{#1}{#2}}}{#2}}
\newcommand{\@@ifnextchar}[3]{\@ifnextchar{#3}{#1}{#2}}
\newcommand{\@@ifnextcharamong}[3]{\@ifnextcharamong{#1}{#2}}
\makeatother

\newcommand{\ucomp}[1]{\hat{#1}}
\newcommand{\finset}[1]{{[#1]}}

\makeatletter
\newcommand{\higherequifibsf}{\mathcal}
\newcommand{\higherequifib}[2]{\higherequifibsf{#1}(#2)}
\newcommand{\underlyinggraph}[1]{U(#1)}
\newcommand{\theequifib}[1]{{\def\higherequifibsf{}#1}}
\makeatother

\newcommand{\loopspace}[2][]{\typefont{\Omega}^{#1}(#2)}
\newcommand{\join}[2]{{#1}*{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Notes on algebraic topology}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\part{Spectral sequences}
\section{Background}
\begin{defn}
A graded $R$-module $M$ is an $R$-module which decomposes as a direct
sum
\begin{equation*}
\bigoplus_{p\in\Z} F_p M
\end{equation*}
of $R$-modules. A graded $R$-homomorphism $h:M\to N$ is an $R$-homomorphism which
decomposes into $h_p:F_pM\to F_pN$. 
\end{defn}

\begin{lem}
Suppose $M$ and $N$ are graded $R$-modules. Then $M\otimes N$ is a graded
$R$-module by
\begin{equation*}
(M\otimes_R N)_i\defeq \bigoplus_{p+q=i} F_pM\otimes_R F_qN.
\end{equation*}
\end{lem}

\begin{defn}
A graded algebra is a graded $R$-module $M$ for which there are linear mappings
$\varphi_{p,q}:F_pM\otimes_R F_qM\to F_{p+q}M$, i.e.~a graded $R$-homomorphism
$\varphi:M\otimes M\to M$, which is associative in the sense
that the diagram
\begin{equation*}
\begin{tikzcd}
M\otimes M\otimes M \arrow[r,"\varphi\otimes 1"] \arrow[d,swap,"1\otimes\varphi"] &
M\otimes M \arrow[d,"\varphi"] \\ M\otimes M \arrow[r,swap,"\varphi"] & M
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{eg}
Polynomials with coefficients in $R$ forms a graded algebra. Moreover, in the
polynomial ring $R[X]$, we find that $G_pR[X]\defeq F_pR[X]/F_{p-1}R[X]\cong R$.
Since those are free modules, we have that $R[X]\cong \bigoplus_p G_pR[X]$. 
\end{eg}

\section{Spectral sequences}

\subsection{Motivation from the long exact sequence of a pair}

Recall that a pair of spaces $(X,A)$ induces a long exact sequence of homology
groups
\begin{equation*}
\begin{tikzcd}
\cdots \arrow[r,"\partial_{n+1}"]
& H_n(A) \arrow[r,"i_n"]
& H_n(X) \arrow[r,"j_n"]
& H_n(X,A) \arrow[r,"\partial_n"]
& H_{n-1}(A) \arrow[r,"i_{n-1}"]
& \cdots
\end{tikzcd}
\end{equation*}
from the short exact sequence
\begin{equation*}
\begin{tikzcd}
0 \arrow[r] & C_\ast(A) \arrow[r] & C_\ast(X) \arrow[r] & C_\ast(X,A) \arrow[r] & 0
\end{tikzcd}
\end{equation*}
of chain complexes, by means of the snake lemma. This long exact sequence helps
us to compute $H_n(X)$ in terms of $H_n(A)$ and $H_n(X,A)$, which may be easier
to determine. For instance, from the long exact sequence we obtain the short
exact sequence
\begin{equation*}
\begin{tikzcd}
0 \arrow[r] & \mathrm{coker}(\partial_{n+1}) \arrow[r] & H_n(X) \arrow[r] & \mathrm{ker}(\partial_n) \arrow[r] & 0
\end{tikzcd}
\end{equation*}
and hence we have determined that $H_n(X)$ can be obtained as some element of the
group $\mathrm{Ext}(\mathrm{coker}(\partial_{n+1}),\mathrm{ker}(\partial_n))$.
In other words, $H_n(X)$ is a particular solution to an extension problem.

Note also that the long exact sequence of relative homology groups can be
presented as an exact triangle of graded $R$-homomorphisms:
\begin{equation*}
\begin{tikzcd}[column sep=0em]
\bigoplus_n H_n(C_\ast(A)) 
\arrow[rr,"i"] & & \bigoplus_n H_n(C_\ast(X)) \arrow[dl,"j"] \\
& \bigoplus_n H_n(C_\ast(X,A)) \arrow[ul,"\partial"]
\end{tikzcd}
\end{equation*}

The first idea of spectral sequences is to generalize the long exact sequence
of homology obtained from a pair of spaces, to an algebraic gadget obtained from
a filtration on a space, and mimic the derivation of determining the homology
group as a solution to an extension problem.

\begin{defn}
A filtration of a space X consists of a sequence 
\begin{equation*}
\cdots\subseteq X_p\subseteq X_{p+1}\subseteq\cdots
\end{equation*}
such that $X=\bigcup_p X_p$ and $\bigcap_p X_p=\varnothing$. A filtration of $X$ is said to be bounded, if
$X_p=\varnothing$ for $p$ sufficiently small, and $X_p=X$ for $X$ sufficiently
large.
\end{defn}

An important class of filtered spaces is that of CW-complexes, where the filtration
$X_p$ of $X$ is given by the $p$-skeleton of $X$. Another case is where
$X_p\defeq\varnothing$ for $p<0$, $X_0\defeq A$ and $X_p\defeq X$ for $p>0$; here
we recover the old theory of the topological pair.

\begin{defn}
Given a space $X$ with a filtration, we can form the staircase diagram
\begin{footnotesize}
\begin{equation*}
\begin{tikzcd}
& \vdots \arrow[d] & & \vdots \arrow[d] \\ 
\cdots \arrow[r]
& H_{n+1}(X_p) \arrow[r] \arrow[d]
& H_{n+1}(X_p,X_{p-1}) \arrow[r]
& H_n(X_{p-1}) \arrow[r] \arrow[d]
& H_n(X_{p-1},X_{p-2}) \arrow[r]
& \cdots \\
\cdots \arrow[r]
& H_{n+1}(X_{p+1}) \arrow[r] \arrow[d]
& H_{n+1}(X_{p+1},X_{p}) \arrow[r]
& H_n(X_{p}) \arrow[r] \arrow[d]
& H_n(X_{p},X_{p-1}) \arrow[r]
& \cdots \\
& \vdots & & \vdots
\end{tikzcd}
\end{equation*}%
\end{footnotesize}%
in which the familiar long exact sequence of the pairs $(X_p,X_{p-1})$ run
down like a staircase.
\end{defn}

\begin{defn}
Let $X$ be a space with a filtration. Then we obtain the exact couple
\begin{equation*}
\begin{tikzcd}
A \arrow[rr,"i"] & & A \arrow[dl,"j"] \\
& E \arrow[ul,"\partial"]
\end{tikzcd}
\end{equation*}
in which $A\defeq\bigoplus_{p,n} H_n(X_p)$, and $E\defeq\bigoplus_{p,n}H_n(X_p,X_{p-1})$. 
\end{defn}

We can come to such an exact couple from any filtered chain complex, which is
what we turn our attention to before continuing.

\subsection{The spectral sequence of a filtered complex}

\begin{defn}
A filtration of an $R$-module $M$ consists of a sequence
\begin{equation*}
\cdots\subseteq F_pM\subseteq F_{p+1}M\subseteq\cdots
\end{equation*}
of $R$-submodules of $M$, such that $M=\bigcup_p F_pM$ and $\bigcap_p F_pM=0$. 
A filtration of $R$ is said to be bounded if $F_pM=0$ for $p$ sufficiently
small and $F_pM=M$ for $p$ sufficiently large.
\end{defn}

\begin{defn}
Let $\{M,F_pM\}$ be a graded $R$-module. The associated graded module is defined
by $G_p M\defeq F_pM/F_{p-1}M$. We obtain a short exact sequence
\begin{equation*}
\begin{tikzcd}
0 \arrow[r] & F_{p-1}M \arrow[r] & F_pM \arrow[r] & G_pM \arrow[r] & 0.
\end{tikzcd}
\end{equation*}
\end{defn}

\begin{rmk}
It would be nice if $F_pM\cong F_{p-1}M\oplus G_pM$, so that we can write
$M\cong\bigoplus_p G_pM$. Under what condition does this hold? This holds if
each $G_pM$ is a projective $R$-module, so under what conditions is this true?
\end{rmk}

\begin{defn}
A filtered chain complex is a chain complex $(C_\ast,\partial)$ together with a
filtration $\{F_pC_i\}$ of each $C_i$, such that the differential preserves the
filtration, i.e.~$\partial(F_pC_i)\subseteq F_p C_{i-1}$. 

A filtration of a chain complex is said to be bounded if it is bounded in each
dimension.
\end{defn}

Let $(F_pC_\ast,\partial)$ be a filtered chain complex. We have again our
short exact sequence
\begin{equation*}
\begin{tikzcd}
0 \arrow[r] & F_{p-1} C_\ast \arrow[r] & F_p C_\ast \arrow[r] & G_p C_\ast \arrow[r] & 0
\end{tikzcd}
\end{equation*}
of chain complexes. This also gives us the long exact sequence on homology,
which we may express conveniently as the exact couple
\begin{equation*}
\begin{tikzcd}[column sep=0em]
\bigoplus_{p,q} H_{p+q}(F_pC_\ast) \arrow[rr,"i"] & & \bigoplus_{p,q} H_{p+q}(F_pC_\ast) \arrow[dl,"j"] \\
& \bigoplus_{p,q} H_{p+q}(G_p C_\ast) \arrow[ul,"k"]
\end{tikzcd}
\end{equation*}
consisting of graded $R$-homomorphisms (of which $k$ shifts in degree).

\begin{defn}
Consider an exact couple, i.e.~a commutative triangle
\begin{equation*}
\begin{tikzcd}
A \arrow[rr,"i"] & & A \arrow[dl,"j"] \\ & E \arrow[ul,"k"]
\end{tikzcd}
\end{equation*}
of $R$-modules, which is exact at every vertex. Taking $\partial^0\defeq j\circ k$,
we see that $(\partial^0)^2=0$ by exactness. We may now form the derived exact couple
\begin{equation*}
\begin{tikzcd}[column sep=0]
\mathrm{im}(i) \arrow[rr,"i'"] & & \mathrm{im}(i) \arrow[dl,"j'"] \\
& \frac{\mathrm{ker}(\partial)}{\mathrm{im}(\partial)} \arrow[ul,"k'"]
\end{tikzcd}
\end{equation*}
where
\begin{align*}
i'(i(a)) & \defeq i(i(a)) \\
j'(i(a)) & \defeq [j(a)] \\
k'([e]) & \defeq k(e)
\end{align*}
\end{defn}

\begin{rmk}
Since quotients commute with direct sums (both are colimits), it follows that
\begin{equation*}
E'\defeq \frac{\mathrm{ker}(\partial)}{\mathrm{im}(\partial)}
  \cong
\bigoplus_{p,q} \frac{\mathrm{ker}(\partial^0_{p,q})}{\mathrm{im}(\partial^0_{p,q+1})}
\end{equation*}
is a graded $R$-module. In other words, $E'$ is a direct sum of the homology
groups of the $p$-indexed family of chain complexes
\begin{equation*}
\begin{tikzcd}
\cdots \arrow[r] & E_{p,q}^0 \arrow[r,"{\partial^0_{p,q}}"] & E_{p,q-1}^0 \arrow[r] & \cdots
\end{tikzcd}
\end{equation*}
It follows that $i'$, $j'$ and $k'$ are graded
whenever $i$, $j$ and $k$ are, where $k'$ shifts down in dimension the same way 
$k$ does.
\end{rmk}

\begin{comment}
\begin{defn}
We define
\begin{equation*}
E_{p,q}^0\defeq G_pC_{p+q}\defeq F_pC_{p+1}/F_{p-1}C_{p+q},
\end{equation*}

Since the differential preserves the filtration, we obtain from the differentials
well-defined $R$-homomorphisms functioning as the boundary maps in the chain complex

\end{defn}

\begin{defn}
The homology groups 
\begin{equation*}
E^1_{p,q}\defeq \mathrm{ker}(\partial^0_{p,q})/\mathrm{im}(\partial^0_{p,q+1})
\end{equation*}
form again a chain complex, with boundary maps $\partial^1_{p,q}:E^1_{p,q}\to
E^1_{p,q-1}$. Thus, this process may be repeated indefinitely.
\end{defn}
\end{comment}

\begin{comment}
\begin{lem}
Let $(C_\ast,\partial)$ be a filtered chain complex. Then there is a filtration
on the homology of $C_\ast$, given by
\begin{equation*}
F_pH_i(C_\ast)\defeq\{\alpha\in H_i(C_\ast)\mid \exists_{(x\in F_p C_i)}\,\alpha=[x]\}.
\end{equation*}
\end{lem}
\end{comment}

\subsection{Convergent spectral sequences}

\begin{defn}
A spectral sequence consists of
\begin{enumerate}
\item An $R$-module $E^r_{p,q}$ for each $p,q\in\Z$ and each $r\geq 0$.
\item Differentials $\partial_r:E^r_{p,q}\to E^r_{p-r,q+r-1}$ such that
$\partial_r^2=0$ and $E^{r+1}$ is the homology of $(E^r,\partial_r)$ 
\end{enumerate}
\end{defn}

\begin{defn}
A spectral sequence $\{E^r,\partial_r\}$ of $R$-modules is said to converge 
if for every $p,q\in\Z$, one has $\partial_r=0:E^r_{p,q}\to E^r_{p-r,q+r-1}$
for $r$ sufficiently large.
\end{defn}

\begin{rmk}
If a spectral sequence $\{E^r,\partial_r\}$ converges, then the $R$-module
$E^r_{p,q}$ is independent of $r$ for sufficiently large $r$. 
\end{rmk}

\begin{thm}
Let $(F_pC_\ast,\partial)$ be a filtered complex. Then we obtain a spectral
sequence $(E^r_{p,q},\partial^r)$ defined for $r\geq 0$, with
\begin{equation*}
E^1_{p,q}\defeq H_{p+q}(G_pC_\ast).
\end{equation*}
This is the spectral sequence of filtered complexes.
\end{thm}

\begin{thm}
If $(F_pC_\ast,\partial)$ is a bounded filtered complex, then the spectral
sequence converges to
\begin{equation*}
E^\infty_{p,q}\defeq G_pH_{p+q}(C_\ast).
\end{equation*}
\end{thm}

Let $X$ be a filtered space, and let our goal be to compute the $n$-th (co)homology
group $H_n(X)$. In general, this might be a complicated task. However, it might
be easier to compute the homologies of the subcomplex $C_\ast(X_p)$, and the quotient
complex $C_\ast(X)/C_\ast(X_p)$. From this, we obtain a short exact sequence
\begin{equation*}
\begin{tikzcd}
0 \arrow[r]
& \mathrm{coker}(\delta) \arrow[r]
& H_\ast(X) \arrow[r]
& \mathrm{ker}(\delta) \arrow[r]
& 0
\end{tikzcd}
\end{equation*}

\subsection{The Serre spectral sequence}

The Serre spectral sequence relates the homology of a Serre fibration to the
homology of the fibers and the base. Thus, in some cases one can compute the
homology of the fibration in terms of the homology of the fibers and the base.

Let $\pi : X\to B$ be a fibration, with $B$ a path-connected CW-complex, and we
filter $X$ by the subspaces $X_p\defeq \pi^{-1}(B_p)$, in which $B_p$ is the
$p$-skeleton of $B$. 

\begin{lem}
The spectral sequence for homology with coefficients in $G$ associated to this
filtration of $X$ converges to $H_\ast(X;G)$.
\end{lem}

\begin{thm}
Let $F\to X\to B$ be a fibration with $B$ path-connected. If $\pi_1(B)$ acts
trivially on $H_\ast(F;G)$, then there is a spectral sequence $\{E^r_{p,q},\partial_r\}$
with:
\begin{enumerate}
%\item $\partial_r : E^r_{p,q}\to E^r_{p-r,q+r-1}$ and $E^{r+1}_{p,q}=\mathrm{ker}\,d_r/\mathrm{im}\,dr$. 
\item the stable terms $E^\infty_{p,n-p}$ are isomorphic to $F^p_n/F^{p-1}_n$ in
a filtration $0\subseteq F^0_n\subseteq\cdots\subseteq F^n_n=H_n(X;G)$ of ...
\item $E^2_{p,q}\cong H_p(B;H_q(F;G))$. 
\end{enumerate}
\end{thm}

\part{K-theory}

\section{Vector bundles}

\subsection{Basic spaces}
\begin{defn}
The \define{$n$-sphere} $\Sn^n$ is the subspace of $\R^{n+1}$ consisting of unit vectors.
The \define{real projective $n$-space} $\R P^n$  is the space of lines in
$\R^{n+1}$ through the origin. Equivalently, we may regard $\R P^n$ as the quotient
space of $\Sn^n$ in which the antipodal pairs of points are identified. Notice
that $\R P^1\approx \Sn^1$. 
\end{defn}

\begin{defn}
For each $n$, we may include the $n$-sphere $\Sn^n$ into $\Sn^{n+1}$ by mapping
it into the equator. These inclusions induce inclusions $\R P^n\to \R P^{n+1}$.
We define $\R P^\infty$ to be the sequential colimit of $\R P^n$. 
\end{defn}

\subsection{Definition and basic properties}
\begin{defn}
An \define{$n$-dimensional vector bundle} is a map $p:E\to B$ together with a
real vector space structure on $p^{-1}(b)$ for each $b\in B$, satisfying the
\define{local triviality condition}, which says that there is an open cover 
$\mathcal{C}$ of $B$, with homeomorphisms $h_U:p^{-1}(U)\to U\times\mathbb{R}^n$
for each $U\in\mathcal{C}$, which maps $p^{-1}(b)$ to $\{b\}\times\mathbb{R}^n$
for each $b\in U$. 

The functions $h_U$ are also called \define{local trivializations}. Given a
vector bundle $p:E\to B$, the space $B$ is called the \define{base space}, the
space $E$ is called the \define{total space}, and the spaces $p^{-1}(b)$ are
called the \define{fibers}. A $1$-dimensional
vector bundle is also called a \define{line bundle}.
\end{defn}

\begin{defn}
An \define{isomorphism of vector bundles} from $p:E\to B$ to $p':E'\to B$
consists of a map $h:E\to E'$ satisfying $p'\circ h=p$, 
which induces a linear isomorphism
$p^{-1}(b)\to p'^{-1}(b)$ between each of the fibers.
\end{defn}

\begin{lem}
If $h:E\to E'$ is an isomorphism of vector bundles, then the underlying map
of type $E\to E'$ is a homeomorphism. 
\end{lem}

\begin{proof}
Suppose $h:E\to E'$ induces isomorphisms $p^{-1}(b)\to p'^{-1}(b)$ for each
$b\in B$. Then, for each $x\in E'$ we have an isomorphism from
$p^{-1}(p'(x))$ to $p'^{-1}(p'(x))$. Since $x\in p'^{-1}(p'(x))$, we find
an element $y\in p^{-1}(p'(x))\subseteq E$. Thus, $h$ is surjective. Now suppose
that $x,x'\in E$ are two elements for which $h(x)=h(x')$. Since $p'\circ h=p$,
it follows that $x'\in p^{-1}(x)$. Now, the fact that $h$ induces an isomorphism
between fibers implies that $x=x'$. 

Thus, $h$ has an inverse function $k:E'\to E$, and we need to show that this
function is continuous. It suffices to show that $k|_U$ is continuous for each
$U$ on which $p'$ is trivial. Let $x\in B$, and compose the map
$h_U:p^{-1}(U)\to p'^{-1}(U)$ with its local trivializations. Thus, we obtain
a map $g_U:U\times\R^n\to U\times \R^n$, mapping $(x,y)$ to $(x,A(y))$, where
$A$ is a linear isomorphism. 
\end{proof}

In the following definition, we give a vector bundle by a gluing construction.

\begin{defn}
Consider a space $B$, and an open cover $\mathcal{C}$ which is closed under
finite intersections. Then $\mathcal{C}$ may be considered a poset ordered by
inclusion.

A \define{collection of gluing functions} consists of a continuous choice of linear 
isomorphisms $g_{U,V}:U\cap V\to GL_n(\R)$ satisfying the \define{cocycle
condition} 
\begin{equation*}
g_{V,W}\circ g_{U,V}=g_{U,W}
\end{equation*}
on $U\cap V\cap W$, for every 
$U,V,W\in\mathcal{C}$. Such a collection of gluing functions determines a functor
$\mathcal{C}\to\mathbf{Top}$, which is given on points by $U\mapsto U\times\R^n$,
and on morphisms by $(x,v)\mapsto(x,A(v))$, for each $U\subseteq V$ determining
a linear isomorphism $A$. 

The colimit of this functor is the total space of a vector bundle.
\end{defn}

\begin{eg}
There are lots of examples of vector bundles:
\begin{enumerate}
\item The \define{$n$-dimensional trivial bundle} over $B$ is defined to be 
$\proj1:B\times\mathbb{R}^n\to B$. So the trivial bundle is the one which is
\emph{globally} trivial. We will write the $n$-dimensional trivial bundle over
$B$ as $\epsilon^n\to B$. 
\item The circle may be regarded as the quotient of $[0,1]$ modulo the end points.
The \define{Mobius bundle} is the line bundle over $\Sn^1$ to have total space
$E\defeq [0,1]\times\R$, with the identifications $(0,t)\sim(1,-t)$. 
\item The \define{tangent bundle} of the unit sphere $\Sn^n$, viewed as a subspace of
$\R^{n+1}$, is defined to be the subspace $E\defeq\{(x,v)\in\Sn^n\times\R^{n+1}
\mid x\perp v\}$ of $\R^{2n+2}$, which projects onto $\Sn^n$. 

The $n$-sphere is covered by $2n+2$ open hemispheres, centering at $\pm e_i$,
where $e_i\in\R^{n+1}$ is a basis vector. 
\item The \define{normal bundle} of the unit sphere $\Sn^n$ is the line bundle
with $E$ consisting of pairs $(x,v)\in\Sn^n\times\R^{n+1}$ such that $v=tx$ for
some $t\in\R$. \emph{The normal bundle on $\Sn^n$ is isomorphic to the trivial line 
bundle $\Sn^n\times\R\to\Sn^n$.}
\item The \define{canonical line bundle} $p:E\to \R P^n$ has as its total space
the subspace $E\subseteq \R P^{n+1}\times\R ^{n+1}$ consisting of pairs
$(l,v)$ with $v\in l$. \emph{The M\"obius line bundle is isomorphic to the
canonical line bundle on $\Sn^1$.}
\item The inclusions $\R P^n\subseteq \R P^{n+1}$ induce inclusions of the
canonical line bundles. The sequential colimit of the canonical line bundles
produces the canonical line bundle on $\R P^\infty$. 
\end{enumerate}
\end{eg}

\begin{defn}
Given two vector bundles $p:E\to B$ and $p':E'\to B$ over the same base space
$B$, we obtain a vector bundle $p\oplus p': E\oplus E'\to B$, fitting in the
pullback square
\begin{equation*}
\begin{tikzcd}
E\oplus E' \arrow[r] \arrow[d] \arrow[dr,"{p\oplus p'}" description ] & E' \arrow[d,"{p'}"] \\
E \arrow[r,swap,"p"] & B
\end{tikzcd}
\end{equation*}
\end{defn}

\begin{eg}
\begin{enumerate}
\item The direct sum of the tangent and normal bundles on $\Sn^n$ is the trivial
bundle $\Sn^n\times\R^{n+1}$. 
\end{enumerate}
\end{eg}

\begin{defn}
Let $p:E\to B$ and $p':E'\to B$ be two vector bundles over the same space $B$,
and choose an open cover $\mathcal{C}$ such that both $E$ and $E'$ are locally
trivial with respect to $\mathcal{C}$. We define $E\otimes E'$ by gluing. 

Then we can define, for each $U,V\in\mathcal{C}$ satisfying $U\subseteq V$, we
have linear isomorphisms $g_{U,V}(x):\R^n\to\R^n$ and $g'_{U,V}(x):\R^m\to\R^m$,
induced by the local trivializations of $E$ and $E'$ respectively. These give
gluing functions $g_{U,V}(x)\otimes g'_{U,V}(x):\R^n\otimes\R^m\to\R^n\otimes\R^m$
for each $x\in U$, and these gluing functions satisfy the cocycle condition.

Thus, we obtain a vector bundle $E\otimes E'$ from these gluing functions.
\end{defn}

\begin{lem}
The tensor product of vector bundles over a fixed base space is commutative,
associative, it has an identity element (the trivial bundle), and it is
distributive with respect to direct sum.
\end{lem}

Change of base $f:B'\to B$ turns a vector bundle $E$ over $B$ to a vector
bundle $f^\ast(E)$ over $B'$.

\begin{lem}
For any two vector bundles $E$ and $E'$ over $B$, and any $f:B'\to B$, we have
natural isomorphisms $f^\ast(E\oplus E')\approx f^\ast(E)\oplus f^\ast(E')$, and
$f^\ast(E\otimes E')\approx f^\ast(E)\otimes f^\ast(E')$. Moreover, if $f$
is homotopic to $g$, then $f^\ast=g^\ast$. 
\end{lem}

\subsection{K-theory}

\begin{defn}
Two vector bundles $E\to B$ and $E'\to B$ are callec \define{stably isomorphic},
if there is an $n$ for which $E\oplus\epsilon^n\approx E'\oplus\epsilon^n$, and
we write $E\approx_s E'$ if $E$ and $E'$ are stably isomorphic. Also,
we will define the relation $E\sim E'$ if there are $m$ and $n$ such that
$E\oplus\epsilon^m\approx E'\oplus^n$. 
\end{defn}

\begin{lem}
The direct sum preserves both $\approx_s$ and $\sim$. Moreover, if $B$ is compact,
then the set of ${\sim}$-equivalence classes of vector bundles forms an abelian
group, called $\tilde{K}(B)$. If $B$ is pointed, then the tensor product turns 
$\tilde{K}(B)$ into a ring.
\end{lem}

\begin{lem}
The direct sum satisfies the cancellation property with respect to $\approx_s$,
i.e.~we have that $E\oplus E'\approx_s E\oplus E''$ implies $E'\oplus E''$. 
Thus, if we define two pairs $(E,F)$ and $(E',F')$ to be equivalent to each
other whenever $E\oplus F'=E'\oplus F$, we obtain an abelian group $K(B)$ for
any compact space $B$. The tensor product turns $K(B)$ into a ring.
\end{lem}

\begin{lem}
We have a ring isomorphism
\begin{equation*}
K(B)\approx \tilde{K}(B)\oplus\Z.
\end{equation*}
\end{lem}

Both $K$ and $\tilde{K}$ are contravariant functors.

\begin{lem}
If $X$ is compact Hausdorff and $A\subseteq X$ is a closed subspace, then the
inclusion and quotient maps $A\stackrel{i}{\to}X\stackrel{q}{\to}X/A$ induces
an sequence
\begin{equation*}
\begin{tikzcd}
\tilde{K}(X/A) \arrow[r,"q^\ast"] & \tilde{K}(X) \arrow[r,"i^\ast"] & \tilde{K}(A)
\end{tikzcd}
\end{equation*}
which is exact at $\tilde{K}(X)$.
\end{lem}

\begin{lem}
If $A$ is contractible, the quotient map $q:X\to X/A$ induces a bijection
$q^\ast:\mathrm{Vect}^n(X/A)\to\mathrm{Vect}^n(X)$.
\end{lem}

Apparently, this gives a long exact sequence of $\tilde{K}$-groups:
\begin{equation*}
\begin{tikzcd}[column sep=small]
\cdots\arrow[r] & \tilde{K}(\Sn(X)) \arrow[r] & \tilde{K}(\Sn(A)) \arrow[r]
& \tilde{K}(X/A) \arrow[r] & \tilde{K}(X) \arrow[r] & \tilde{K}(A).
\end{tikzcd}
\end{equation*}
Still considering pointed spaces, we may consider the long exact sequence of the pair $(X\times Y,
X\vee Y)$. Recall that $(X\times Y)/(X\vee Y)$ is the smash product
$X\wedge Y$, i.e.~the smash product is the pushout of $\unit\leftarrow
X\vee Y\rightarrow X\times Y$. The long exact sequence of the pair
$(X\times Y,X\vee Y)$ looks as follows:
\begin{equation*}
\begin{tikzcd}[column sep=.8em]
\cdots\arrow[r] & \tilde{K}(\Sn(X\times Y)) \arrow[r] & \tilde{K}(\Sn(X\vee Y)) \arrow[r]
& \tilde{K}(X\wedge Y) \arrow[r] & \tilde{K}(X\times Y) \arrow[r] & \tilde{K}(X\vee Y).
\end{tikzcd}
\end{equation*}  

\subsection{Bott periodicity}

\begin{defn}
We define an \define{external product} $\mu:K(X)\otimes K(Y)\to K(X\times Y)$,
by $\mu(a\otimes b)\defeq \proj1^\ast(a)\cdot\proj2^\ast(b)$. 
\end{defn}


\end{document}
